### 2014-04-09 paper/discussion.tex
final sections

The promise of 3D printing is two-fold: it enables the manufacturing of complex objects at low volumes (n=1), and it allows for the manufacturing of geometries that are impossible to create using traditional techniques like injection molding.  Printing pipes to prototype interactive objects exploit both advantages. Prototypes by definition are low-volume parts, and pipes require complex interior geometries that are very difficult or impossible to produce using other processes. Our work has focused on routing pipes that are then later manually filled after a print has completed. However, the fundamental operations of 3D path routing and topology specification that our tool offers will remain relevant as 3D printing technology advances. Multi-material printers are already available; as they become able to lay down additional materials (e.g., conductive), \systemname can enable designers to efficiently specify where different functional materials should be placed at design time.

@@

%Some 3D printers can fabricate rubber-like materials.  Similar to Slyper, et al., in \cite{Slyper-shape}, we can sense \emph{flexing} and \emph{bending} of prints made on these machines: the ability to 3D print devices to sensing bending and flexing significantly cuts molding and assembly time.

\emph{Tapping and Grasping} can be sensed acoustically.  Due to differential sound conductivity in different printed materials, solid pipes of a material with greater conductivity (i.e., rigid, high-density) can be embedded in a model of a lesser conductivity (i.e., softer).  A microphone or piezo placed at the system-side of these sound-conducting pipes can determine where the model was tapped and how it was tapped.  Active acoustic sensing as described in Touch \& Activate \cite{Ono-touchandactivate} is also possible; this technique would allow for particular areas of interest (connected to rigid pipes) to be touch sensitive and enable grasp sensing.

\subsection{Identification}
\bjoern{this section should be about half as long as it is right now.}
In addition to active inputs from users, pipes can be used for passive \emph{identification}.  While the pipes need not be visible (especially in the case of fully-enclosed pipes), their presence, location, network structure, and length change the physical properties of a printed object.  This includes weight, acoustic resonance, and capacitive signature.b

Both identification by recall and intentional encoding are possible.  As seen in Touch \& Activate \cite{Ono-touchandactivate}, different objects have different acoustic signatures.  Additionally, two objects that are visually identical but which have fully enclosed (or other types) of pipes on their interior can be distinguished acoustically.  Thus we can recall an object's identity once its acoustic signature has been recorded.  Similarly, the Touch\'{e} system \cite{Sato-touche} relies on distinguishing capacitive signatures of objects, both in static states and as their boundary conditions are changed by human interaction.  Weight is a simple metric of object identity, but weights are less distinctive than acoustic or capacitive signatures, as they are represented by a single number. \bjoern{second half of this paragraph meanders. focus on the key story.}

We have experimented with intentional encoding.  Semi-closed pipes can function as resonance chambers.  A pipe's resonant frequency can be measured by attaching a speaker and microphone to it, sweeping frequencies with the speaker, and performing a Fourier transform on the resultant signal from the microphone.  Peaks in the transformed data correspond to resonant frequencies of the object.  Intentional encoding in the case of capacitance is significantly more difficult, and we leave it to future work.  Encoding of weight is trivial by creating fully-enclosed chambers inside an object.  This is similar to Infrastructs \cite{Willis-infrastructs}.  While their technique requires access a terahertz imaging tool, audio resonance identity encoding requires only a microphone and a speaker.

@@

\subsection{Printed Conductors and Passive Components}

Sensing via traditional electronic components (e.g., p botentiometers, alcohol gas sensors) can be accomplished by conductive inserted media.  Components can be recessed into a print's surface, with their leads implanted into open pipes.  \bjoern{didn't we already do this in the radio?} By inserting liquid copper paint instead of threading traditional wires, all components in a model can share a ground line, and the paint's drying process obviates the use of solder or glue to affix the components in place.

>>>>

The promise of 3D printing is two-fold: it enables the manufacturing of complex objects at low volumes, and it allows for the manufacturing of geometries that are impossible to create using traditional techniques like injection molding.  Printing pipes to prototype interactive objects exploit both advantages. Prototypes by definition are low-volume parts, and pipes require complex interior geometries that are very difficult or impossible to produce using other processes. Our work has focused on routing pipes that are then later manually filled after a print has completed. However, the fundamental operations of 3D path routing and topology specification that our tool offers will remain relevant as 3D printing technology advances. Multi-material printers are already available; as they advance and support the embedding of additional materials (e.g., conductive), \systemname can enable designers to efficiently specify where different functional materials should be placed at design time.

@@

%Some 3D printers can fabricate rubber-like materials.  Similar to Slyper, et al., in \cite{Slyper-shape}, we can sense \emph{flexing} and \emph{bending} of prints made on these machines: the ability to 3D print devices to sensing bending and flexing significantly cuts molding and assembly time.
As demonstrated in our Custom Radio example object, most forms of input can be easily supported by wiring traditional electronic compents through open pipes that connect to a microcotnroller and power source. Components can be recessed into a print's surface, with their leads inserted into the open pipes.  By inserting liquid copper paint instead of threading traditional wires, all components in a model can share a ground line, and the paint's drying process obviates the use of solder or glue to affix the components in place. We have also shown how touch input can be enabled using conductive paint and SFCS \cite{Sato-touche}.

Alternatively, it may be possible to sense tapping and grasping acoustically.  Due to differential sound conductivity in different printed materials, solid pipes of a material with greater conductivity (i.e., rigid, high-density) can be embedded in a model of a lesser conductivity (i.e., softer).  A microphone or piezo placed at the system-side of these sound-conducting pipes can determine where the model was tapped and how it was tapped.  Active acoustic sensing as described in Touch \& Activate \cite{Ono-touchandactivate} is also possible; this technique would allow for particular areas of interest (connected to rigid pipes) to be touch sensitive and enable grasp sensing.

\subsection{Identification}
\bjoern{this section should be about half as long as it is right now.}
In addition to active inputs from users, pipes can be used for passive \emph{identification}.  Fully-enclosed pipes which ar enot visually apparent can change that weight, acoustic resonance, and capacitive signature of a 3D object.

Each of these properties could be used for object identification. As seen in Touch \& Activate \cite{Ono-touchandactivate}, different objects have different acoustic signatures.  As such, two objects that are visually identical but have differing internal chamber geometry on their interior may be distinguished acoustically.  Thus we can recall an object's identity once its acoustic signature has been recorded.  Similarly, the Touch\'{e} system \cite{Sato-touche} relies on distinguishing capacitive signatures of objects, both in static states and as their boundary conditions are changed by human interaction. Weight could also be used as a simple metric of object identity, but would be less distinctive than acoustic or capacitive signatures.

%We have experimented with intentional encoding.  Semi-closed pipes can function as resonance chambers.  A pipe's resonant frequency can be measured by attaching a speaker and microphone to it, sweeping frequencies with the speaker, and performing a Fourier transform on the resultant signal from the microphone.  Peaks in the transformed data correspond to resonant frequencies of the object.  Intentional encoding in the case of capacitance is significantly more difficult, and we leave it to future work.  Encoding of weight is trivial by creating fully-enclosed chambers inside an object.  This is similar to Infrastructs \cite{Willis-infrastructs}.  While their technique requires access a terahertz imaging tool, audio resonance identity encoding requires only a microphone and a speaker.

@@

\subsection{Passive Components}

### 2014-04-09 paper/exampleobjects.tex
final sections

\section{Example Objects}
\bjoern{subsection headings need to be updated in light of the simplification and label changes in the design space.}
To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using \systemnamenospace.  All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

@@

The models we used for the toys' exteriors were all downloaded from the free website Thingiverse.  All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running open-source SFCS code\footnote{http://www.instructables.com/id/Touche-for-Arduino-Advanced-touch-sensing}.

@@

To demonstrate how gases can be used to both provide sensation to the user through openings and to deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:breathe}).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump: one terminal creates a vacuum while the other creates positive pressure.  Our rabbit has two pipes, one open tube exiting at its mouth and one semi-enclosed, capped pipe capped in its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on our Objet in rubberlike material, with support flushed post-print.

\subsection{Custom Radio (open, threadable, terminals)}

A custom radio shows how pipes can be used to integrate electronic components into the user-facing surface of an object. This device uses a network of disconnected open pipes threaded with wires to connect our components (two potentiometers, an LED, and a speaker) to a microcontroller (Arduino Pro Micro). We placed the microcontroller, an Si4703 radio tuner breakout board and a battery-driven power supply in the base of the radio.

This design with 8 pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) showed the limitations of our incremental routing strategy: since we voxelize and remesh for each pipe, surface detail is incrementally lost. The problem can be addressed by increasing the voxel resolution. At 512x512x512 voxels, surface geometry is not notably degraded. However, a single cut operation takes around 3 minutes \bjoern{on what hardware? mention earlier}, compared to near-instant performance on a $128^3$ voxel grid.  We fabricated the radio itself on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted.

\bjoern{I ran out of steam somewhere here...}
\subsection{Presence-aware Pen Holder (open, solid, ternimals)}
\bjoern{isn't this a threadable instead of a solid?} \bjoern{the model looks like the acutal pipes are really short and basically straight at the bottom of the model. This may raise the question if one couldn't just use the line sensors without any fiber optic cable in the middle.}
Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:pens}).  Our pen holder uses a modification of the FlyEye technique described by Wimmer in \cite{Wimmer-flyeye} in which \bjoern{describe...}. Four internal pipes are threaded with fiber optic cables, one per pen chamber.  We use a single cable per chamber; our 6mm diameter fiber optic cables, in comparison to the fine tubes used in the original work , can send and receive through the same cable. \bjoern{is this because it has many strands?}  At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.  This prototype was built by our Objet and support was flushed post-print.\bjoern{Incorrect. Looks like a Makerbot print.} \valkyrie{this whole paragraph is unclear}

@@

Neon art, perhaps best known for its association with Las Vegas, is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  The pipes have been threaded with electroluminescent wire which is lit in sequence to create an animation.  Our sign was fabricated on the Objet in two halves, cutting through the plane of pipe path to facilitate assemblhy. Pipes enter and exit through the rear to hide the EL wire controls.  The EL wire was threaded post-print, and is controlled by an EL wire sequencer.

@@

Using a modification of our graph-based algorithm which only cuts the literal drawing (and does not perform Eulerization, etc.), we created a maze game with a single particle trapped in a fully enclosed tube.  This maze is based on an SVG we created by hand and processed with our internal path tube tool.  To fabricate the tube, we created it in two halves that were fastened together via glue, due to limitations in our ability to remove support material otherwise (we only have access to a high-pressure water tool, not a lye bath).  We created this prototype on our Objet in two halves, inserting a marble before closing. \bjoern{check - i think the latest maze is in fact done with our tool? Will need rewriting.}

>>>>

\section{Example Objects}
\tovi{it might be intersting to relate our examples to existing HCI papers - in order we have: Touche, Dynamically Changeable Physical Buttons/PneUI, Midas, Suaron. I guess the maze doesnt really relate to anything.}
\tovi{UIST is someitmes cranky about providing enough details so other can reporduce the work. We might want to be more careful about providing specific part numbers of specifications.}
\bjoern{subsection headings need to be updated in light of the simplification and label changes in the design space.}
To evaluate and highlight our tool's capabilities, we fabricated a set of five prototype objects designed using \systemnamenospace.  \tovi{might want to specify the exact printer models we used here, and say we used multipel printers to show that are approach is printer independant.}
%All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

@@

The models we used for the toys were downloaded from the  Thingiverse website. We then used \systemnamenospace to create the desired internal pipes. All components were fabricated on a Makerbot, support-free.  After printing, we injected copper paint (CuPro-Cote) into the interior pipes using a craft syringe.  This paint requires approximately 2 days to fully dry in this configuration (3mm diameter tubes, maximum tube length 10cm).  Our smart base is powered by an Arduino Uno running open-source SFCS code\footnote{http://www.instructables.com/id/Touche-for-Arduino-Advanced-touch-sensing}. An advatnage of using this paint is that it also serves as an adhesive, so no soldering was required. \tovi{confirm}

@@

To demonstrate how gases can be used to both provide sensation to the user through openings and to deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:breathe}).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump: one terminal creates a vacuum while the other creates positive pressure.  
The rabbit model was downloaded from the Thingiverse website. \tovi{confirm} We then used \systemnamenospace to add two pipes: one open pipe exiting at its mouth, and one semi-closed pipe capped at its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on the Objet using a flexible material and its support material was flushed post-print.\tovi{give pump part number/details}

\subsection{Custom Radio (open, threadable, terminals)}
\tovi{stick with the strucutre - describe what it does, then describe modeling and fabrication (how it was built)}
A custom radio shows how pipes can be used to integrate electronic components into the user-facing surface of an object. The radio includes a power and volume dial, and LED which goes on when the radio is on, and a tuner dial. \tovi{confirm}. A Arduino Pro Micro microcontroller, Si4703 radio tuner, and  [x volt] power supply are embedded in the base of the radio. The sound is played through a single speaker.

The radio was manually modelled in SolidWorks. We included slots on the 3D model matching the shape of the electronic components that were to be embedded. The model was then imported into \systemnamenospace, where we created a network of 8 individual open pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) to connect the the comments to the base of the radio. We fabricated the radio itself on our Makerbot, support-free, but cut in two pieces to allow the Arduino and battery pack to be inserted. We then threaded the pipes with wires to connect the compoenents to the microcontroller at the base of the radio.

%This design with 8 pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) showed the limitations of our incremental routing strategy: since we voxelize and remesh for each pipe, surface detail is incrementally lost. The problem can be addressed by increasing the voxel resolution. At 512x512x512 voxels, surface geometry is not notably degraded. However, a single cut operation takes around 3 minutes \bjoern{on what hardware? mention earlier}, compared to near-instant performance on a $128^3$ voxel grid.  We fabricated the radio itself on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted. TOVI: Move to discussion if you want to keep this text.

\bjoern{I ran out of steam somewhere here...}
\tovi{relate this to sauron as well}
\subsection{Presence-aware Pen Holder (open, solid, ternimals)}
\bjoern{isn't this a threadable instead of a solid?} \bjoern{the model looks like the acutal pipes are really short and basically straight at the bottom of the model. This may raise the question if one couldn't just use the line sensors without any fiber optic cable in the middle.}
Our presence-aware pen holder can distinguish which tool or tools a user has picked up (see Figure \ref{fig:pens}).  Our pen holder uses a modification of the FlyEye technique described by Wimmer in \cite{Wimmer-flyeye} in which \bjoern{describe...}.

The model was created/downloaded from [x]. We used \systemnamenospace to add four internal open pipes connecting each pen chamber to the based on the pen holder. This prototype was built by our Objet and support was flushed post-print.\bjoern{Incorrect. Looks like a Makerbot print.} \valkyrie{this whole paragraph is unclear}

Each pipe was then threaded with a fiber optic cable.  We use a single cable per pipe; our 6mm diameter fiber optic cables, in comparison to the fine tubes used in the original work , can send and receive through the same cable. \bjoern{is this because it has many strands?}  At the base of each tube is a QRE1113 line sensor digital breakout board, which has an integrated IR emitter and receiver.   When a pen is in its appointed place, the emitted infrared light is reflected off its bottom and travels back to the receiver, where it registers as bright.

@@

Neon art is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''.  
\tovi{in the system seciton do we ever discuss how entry and exit points are specified for path-constrained pipes?}
The geomertry of the model was imported into \systemnamenospace, which created the assocaited Pipes using its internal path routing tool. We then printed the model on the Objet using a transparent material. The model was printed in two halves, cutting through the plane of pipe path, to facilitate assembly.

The pipes were threaded with electroluminescent wire which is lit in sequence to create an animation using an EL wire sequencer. Pipes enter and exit through the rear to hide the EL wire controls.

@@

Our final example object is a maze game with a marble that can be navigated through a maze-like path, with exits on either side of the model.

The maze is based on an SVG file we manually created and processed with our internal path routing tool.  The Maze was printed on the Objet using a transparent material. We created the maze in two halves that were fastened together via glue, to support the removal of support materials using a high-pressure water tool.  \bjoern{check - i think the latest maze is in fact done with our tool? Will need rewriting.}
